### 2014-11-18 poster/poster.tex
draft 1 of poster

%==============================================================================
%== template for LATEX poster =================================================
%==============================================================================
%
%--A0 beamer slide-------------------------------------------------------------
\documentclass[final]{beamer}
\usepackage[orientation=portrait,size=a0,
            scale=1.25         % font scale factor
           ]{beamerposter}

\geometry{
  hmargin=2.5cm, % little modification of margins
}

%
\usepackage[utf8]{inputenc}

\linespread{1.15}
%
%==The poster style============================================================
\usetheme{sharelatex}

%==Title, date and authors of the poster=======================================
\title
[CSSS 501 MLE Poster Session, 17 November 2014 -- http://github.com/pschmied/bikecounts-mle] % Conference
{ % Poster title
  Weather and Seasonal Factors Affecting Daily Bicycle Counts at
  Fremont St. Bridge, Seattle %
}

\author{ % Authors
Peter Schmiedeskamp\inst{1}, Weiran Zhao\inst{1}
}
\institute
[University of Washington] % General University
{
\inst{1} University of Washington, Seattle
}
\date{\today}

\begin{document}
\begin{frame}[t]
%==============================================================================
\begin{multicols}{3}
%==============================================================================
%==The poster content==========================================================
%==============================================================================

\section{Introduction \& Research Questions}
This project develops a statistical model for predicting the volume of
bicycle traffic per day on the Fremont bridge, in order to better
understand factors believed to influence likelihood of bicycling.
These predictors include weather factors such as precipitation and
temperature, season, day of the week, as well as consideration of a
general trend over time.

\section{Data}
In late 2012, the Seattle Department of Transportation deployed an
automated bicycle detection system that has been collecting bicycle
counts since its installation. These data were downloaded from the
City of Seattle's public data portal, aggregated by day, and then
joined by date to historic weather data queried from Forecast.io. In
this analysis, the collection period starts on October 31, 2012 and
ends October 30, 2014, for a total of 730 daily observations.

These data are summarized graphically in figures \ref{fg:hist} and
\ref{fg:timeseries} below. From the histogram, we see what looks to be
a Poisson distribution with over dispersion. From the timeseries plot,
we see that there is a clear yearly pattern, as well as a weekly
pattern.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/dp1}
\caption{Histogram of bicycle counts}
\label{fg:hist}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/dp2}
\caption{Counts by day}
\label{fg:timeseries}
\end{center}
\end{figure}

\section{Model Specification}
Because over dispersion was evident in the counts, we chose to fit a
negative binomial model. This model includes the following variables,
and was fit using the \texttt{pscl::glm.nb()} function in R.

\begin{table}[htdp]
\begin{center}
\begin{tabular}{ll}
  \hline\hline
  Variable & Description \\
  \hline
  Count* & Number of bicycles / day \\
  Daylight hours & Length of daylight in hours \\
  Holiday & Was the day a holiday (recognized by UW)? \\
  UW & Was University of Washington in session? \\
  Max temp & Maximum temperature for the day \\
  Max precip & Daily max precip (inches) in any hour \\
  Day number & Sequentially numbered day of study \\
  Sat--Fri & Dummy variables for day of week \\
  \hline\hline
    & \multicolumn{1}{r}{*dependent variable}
\end{tabular}
\end{center}
\label{default}
\end{table}%

\section{Model Fit and Diagnostics}

As the actual-versus-predicted plot below shows, the model is a
reasonably good fit. However, we note several outliers falling far
from the $y=x$ line, perhaps indicating some other unaccounted source
of variation in the dataset.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.15\textwidth]{../fig/avpp1}
\caption{Actual versus predicted values from model}
\label{fg:avp}
\end{center}
\end{figure}

\section{Simulated Counterfactual Results}

\subsection{Temperature}

As expected, the effect of temperature seems clear. Bicyclists at this
location appear to prefer warmer weather universally over cooler
weather. In a less temperate climate, we might expect bicycle counts
to fall again in very high temperature conditions.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/m1c1}
\caption{Simulated effect of daily maximum temperature on bicycle counts}
\label{fg:temp}
\end{center}
\end{figure}

\subsection{Precipitation}
Similiarly, we see the strong negative effect of precipitation in our simulated
results.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/m1c2}
\caption{Simulated effect of daily precipitation (in inches) on
  bicycle counts}
\label{fg:maxtemp}
\end{center}
\end{figure}

\subsection{Day of the Week}
As identified in our descriptive timeseries plot, these simulated
results confirm the presense of a weekly pattern in bicycling, with
much higher numbers of bicyclists during the week, falling off
somewhat toward Friday, and dropping precipitously on the weekends.
\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/m1c3}
\caption{Simulated bicycle counts differences by day of week}
\label{fg:dow}
\end{center}
\end{figure}

\subsection{Seasonality}
In considering how to address seasonality, we were concerned that
Seattle's Pacific Maritime Climate does not reflect a traditional four
season calendar year. We felt that hours of sunshine and whether or
not it was the ``school season'' (i.e. the University of Washington
was in session) were better approximations of season for Seattle.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/m1c4}
\caption{Effect of season (measured continuously as number of daylight
  hours)}
\label{fg:daylight}
\end{center}
\end{figure}

\subsection{General trend over time}
In our model we included a term that was a sequential numbering of the
days of the study period as to model a general trend over time. In the
visualization of the simulated results, we do see a general increase
in bicycle counts over this two year span.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/m1c5}
\caption{Simulated bicycle counts over study period}
\label{fg:trend}
\end{center}
\end{figure}

\section{Conclusions}
Over all, this model appears to be reasonably
predictive, with the results confirming our intuitions about factors
influencing bicycling rates. One major limitation, is that we still do
not account for the possibility that there is some timeseries
autoregression ocurring on a day-to-day basis. This is a problematic
limitation, in that we believe that the likelihood of bicycling one
day is very likely influenced by the choice to bicycle on the days
immediately preceeding.

%==============================================================================
%==End of content==============================================================
%==============================================================================

%--References------------------------------------------------------------------

% \subsection{References}

% \begin{thebibliography}{99}

% \bibitem{ref1} J.~Doe, Article name, \textit{Phys. Rev. Lett.}

% \bibitem{ref2} J.~Doe, J. Smith, Other article name, \textit{Phys. Rev. Lett.}

% \bibitem{web} \url{http://www.google.pl}

% \end{thebibliography}
%--End of references-----------------------------------------------------------

\end{multicols}

%==============================================================================
\end{frame}
\end{document}