### 2014-12-25 article/article.tex
final rev for class

% !TeX TS-program = pdflatexmk
% !BIB TS-program = biber

\documentclass[12pt,letterpaper,article,twocolumn]{memoir}

% Layout options
\chapterstyle{article}
\setulmarginsandblock{.75in}{1in}{*}
\setlrmarginsandblock{.75in}{*}{*}
\setcolsepandrule{.33in}{0pt}

\checkandfixthelayout
\raggedbottom

% Bibliography
\usepackage[authordate,strict,bibencoding=inputenc,doi=false,isbn=false,annotation=true]{biblatex-chicago}

% Graphics
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\addbibresource{../bibliography/bikecounts.bib}

% Float barriers
\usepackage[section]{placeins}

\title{Estimating Daily Bicycle Counts in Seattle from Seasonal and Weather Factors}
\author{Peter Schmiedeskamp, Weiran Zhao}
%\date{8 October 2014} % Delete this line to display the current date

%%% BEGIN DOCUMENT
\begin{document}

\maketitle

\section*{Introduction}
%
% 1. Explain what we set out to do (i.e. understand influence of
% seasonal and weather factors on bicycling) 2. explain relevance to
% planners (e.g. understanding how much seasonal variability to expect
% for service provision, understand methodologically how to use data
% collected from counters) 3. ??? other things I'm forgetting right
% now.
%

Bicycling offers a variety of social benefits that include the health
of those choosing to bicycle regularly, as well as, more broadly, a
society-wide increase in mobility when people shift from congestion
causing modes to bicycling. As such, bicycling is being promoted by
policy makers and urban planners as an increasingly important
component of urban transportation. In order to develop policy and
infrastructural improvements to induce more bicycling, a robust
understanding of factors related to bicycling is required.

This article contributes to that understanding by focusing on the
substantive effects of weather and seasonality on the number of
bicyclists observed on a given day. Specifically, we hypothesize that
factors associated with inclement weather (esp. low temperature and
precipitation) will be inversely correlated with bicycle counts. We
similarly expect to see an additional decrease in counts associated
with both physical and social/institutional dimensions of season. That
is, in seasons where the days get shorter, we expect fewer cyclists
due to a kind of decision making inertia. We also expect fewer
cyclists outside the ``school season'' as we believe that a large
number of bicycling trips are associated with commuting to schools
that, unlike most other institutions have clear annual cycles.

In order to address these two primary research questions, we develop a
statistical model estimated using two years of automated bicycle
counts collected on Seattle's Fremont Street bridge, as well as
historic weather and season data collected from other sources.

Using this model, we were able to identify and quantify substantial
effects on bicycle counts associated with temperature, precipitation,
hours of daylight, and school session status. Additionally, we
estimated changes in bicycle numbers associated with the day of the
week and holidays, as well as an overall trend toward increased
bicycling.

\section*{Literature review}
% 
% Identify where people have talked about weather and seasonal factors
% influencing counts previously. Describe their methods together with any
% limitations, as well as their substantive findings---especially what
% variables they considered. Normally should conduct this first, but
% our driver here is a class assignment! We should fill this in and
% adjust our analysis accordingly before attempting to go to press.
%

Bicycling volume in cities is useful for practitioners and researchers
to understand safety, travel behavior, and development impacts.
Therefore the relationship between bicycle volume and various factors,
with the goal to build a predictive model based on this relationship,
has been of great interest to researchers over the last decade (e.g.
\cite{Griswold:2011aa,Fields:2012aa,Niemeier:1996aa,Nosal:2014aa}).

To address these questions, researchers have used a variety of
approaches using different data sources. Studies employing survey or
census data are more often used to explain influences such as
demographic and socioeconomic factors on mode choice
\parencite{Parkin:2008aa,Helbich:2014aa}. In other studies, collected
count data (from automated counters or manual counts) have more been
used recently to track and analyze counts over longer periods of
time \parencite{Griswold:2011aa,Nosal:2014aa}. These counts-based
approaches have largely followed a similar approach of first proposing
a set of explanatory variables, fitting some form of regression model,
and then interpreting / justifying the results vis-\`{a}-vis the
guiding theory.

A literature review accompanying a recent report by
\textcite{Bassok:2011aa} identified eleven primary indicators. These included 
time of day \parencite{Schwartz:1999aa}, season \parencite{Niemeier:1996aa},
population and employment densities
\parencite{McCahil:2008aa,Pinjari:2009aa}, land-use mix
\parencite{Pinjari:2009aa}, bicycle facility type \parencite{Hunt:2007aa},
traffic volume \parencite{McDonald:2007aa}, rain and temperature
\parencite{Niemeier:1996aa,Parkin:2008aa},
income \parencite{Turner:1998aa}, and age \parencite{Hunt:2007aa}.
This section outlines the key points made in the literature that are
relevant to some of most important variables.

Research has found the variability for counts has a positive
association with high temperature and low precipitation. \parencite{Niemeier:1996aa,Parkin:2008aa}. Meanwhile, as suggested by
\textcite{Lewin:2011aa} and \textcite{Thomas:2009aa}, the effects of
precipitation and temperature on bicycle volumes are nonlinear. For
example, bicycle traffic can decrease in both very cold and very hot
weather as noted by \textcite{Richardson:2000aa}. Apart from the usual
temperature and rain variables, \textcite{Miranda-Moreno:2011aa} finds
humidity and additional precipitation variables including the presence
of rain in the morning and/or during the previous three hours to be
significant too. Other comparative studies are also available where
bicycle counts are conducted in different cities, and different
sensitives to weather are examined \parencite{Rose:2011aa}. As for
longitudinal studies, \textcite{Niemeier:1996aa} finds increased variability
for counts conducted in the later months of the year.
\textcite{Jones:2010aa} conclude that morning peak hours from 6 AM to 9 PM
accounts for a consistent 95\% of the total bicycle volumes by hourly
count data. The simple linear regression model has been used in
several applications \parencite{Jones:2008aa,Jones:2010aa}. Other modeling approaches include \textcite{Miranda-Moreno:2011aa} which develops a count
model and \textcite{Thomas:2009aa} which develops a time-series model.
\textcite{Niemeier:1996aa} also uses a Poisson model to statistically
confirm  many of the factors thought to influence cyclists. The
work by \textcite{Gallop:2012aa} adopts a similar time-series approach
while incorporating an autoregressive integrated moving average
(ARIMA) analysis. A summary of findings from the literature are
presented in table \ref{tb:lit}.

These literature together suggest an opportunity for further model
development base around long-term automated counts utilizing
appropriate statistical methods. How seasonal factors influence
bicycle flow needs to be examined in data that last more than a year.
One limitation present in much of the past literature is that few
discuss goodness of fit of their modeling. Further a model that can
better describe and forecast the bicycle count in longitudinal form is
necessary to developed. Models for count data with better estimation
methods offer some promise.

\newfixedcaption{\freetabcaption}{table}
\begin{centering}
\begin{scriptsize}
\freetabcaption{Summary of findings from literature review}
\begin{ctabular}{p{.6in}p{1.15in}p{1.15in}}
  \toprule
  Source & Variable(s) identified & Methods \\
  \midrule
  \cite{Fields:2012aa} & Average daily temperature (F); Total weekly precipitation (inches) & Identify patterns through scatter plots; No explicit model is established. \\
  \cite{Gallop:2012aa} & Temperature, Relative humidity, wind speed, clearness,fog, precipitation & Use ARIMA to account for serial correlation patterns \\
  \cite{Griswold:2011aa} & Nearby population and employment density, proximity to downtown/freeway, age, education level, income, et. & Log linear ordinary least squares regression is used to estimate a bicycle count model \\
  \cite{Helbich:2014aa} & Daily maiximum air temperature (C), daily average wind speed (m/s), daily precipitation(mm) & Place-specific associations of weather conditions are explored through geographically weighted logit models \\
  \cite{Hunt:2007aa} & Descriptive variables indicating lane use, secured parking, level of experience, etc. & Logistic model of cycling-related choices \\
  \cite{Jones:2010aa} & Length of bicycle network, employment density, population density & Standard ordinary least squares regression \\
  \cite{Lewin:2011aa} & Max temp (F), rain flag, snow flag, weekend flag, over 90 flag & Standard linear regression model \\
  \cite{McCahil:2008aa} & logarithmic choice measure, population density, worker density & A new space syntax theory is used to evaluate and predict the bicycle volume throughout a network \\
  \cite{Miranda-Moreno:2011aa} & Temperature (C), humidity (\%), rain presence (0/1), rain in prev. 3hrs (0/1), warm \& humid (0/1), morning rain (0/1) & Both log-linear model and negative binomial model are tested \\
  \cite{Niemeier:1996aa} & Morning flag, rain flag, high temp flag, location variable, season variable & A Poisson bicycle count model is assumed and fitted \\
  \cite{Nosal:2014aa} & Temperature (F), humidity (\%), rain flag, rain prev. 3hrs, am rain, pm rain & The relationship  is analyzed using a log-linear regression model \\
  \cite{Parkin:2008aa} & Gender, car ownership, hilliness, off-road routes proportion & A logistic regression model of relevant socio-economic and physical variables is estimated. \\
  \cite{Pinjari:2009aa} & Household density, employment density, fraction of commercial land area, fraction of senior, fraction of race population, etc. & The model system takes the form of a joint mixed Multinomial Logit–Multiple Discrete-Continuous Extreme Value (MNL-MDCEV) structure \\
  \cite{Rose:2011aa} & Temerature(C), rainfall(mm), holiday flag, school seanson flag, day of the week & Weather and other effects examined using an aggregate model of daily ridership \\
  \cite{Thomas:2009aa} & Temperature(C), sunshine(h), precipitation(mm), wind force(m/s), cycle path use & A bi-level structure is developed with the upper level being a log-linear model and the lower level being a linear model \\
  \bottomrule
\label{tb:lit}
\end{ctabular}
\end{scriptsize}
\end{centering}

\section*{Methodology}
In order to discern the relationship between bicycle counts and
several identified weather, seasonal, and temporal factors, we
developed a statistical model that attempts to predict daily bicycle
counts from these other factors. This section describes the methods
and procedures we used to collect and process the estimation dataset,
the rationale for our selection of variables, our chosen model type,
and our model estimation procedures.

\FloatBarrier
\subsection*{Data collection, processing, description}
%
% Data collection: study site, descriptives of sample, collection from
% forecast.io and seattle.gov, not in that order
%

The data that we attempt to explain were collected at Seattle's
Fremont Bridge and cover a period of two years spanning from October
31, 2012 to October 30, 2014. The Fremont Bridge captures a
substantial amount of bicycle traffic due to its status as one of only
five facilities that carry bicyclists across the canal separating the
northern and southern halves of Seattle.

Bicycle counts were collected at this location continuously by the
City of Seattle using an in-sidewalk counter manufactured by
EcoCounter. When a bicycle passes over an induction loop embedded in
the sidewalk on either side of the Fremont Bridge, the counter
registers the bicycle. Bicyclists may legally choose to ride in the
roadway instead of the sidewalk, and would thus not be detected by the
counter. However, we believe these crossings are rare at this location
due to the design of the facility, which directs bicyclists to enter
the sidewalk, and from our own experience riding and observing other
riders. Counts are aggregated into 15 minute intervals by the City of
Seattle, and are made available to the public via the City of
Seattle's data
portal \parencite{City-of-Seattle:aa,City-of-Seattle:ab}.

Weather data are collected by a variety of sources and are aggregated
by Forecast.io. These data are available through the company's web
services API \parencite{The-Dark-Sky-Company:aa}. Historical daily
summaries are available for a range of weather variables including
several specifically important to our model such as precipitation,
daily minimum and maximum temperatures, sunrise, and sunset.

We downloaded and processed these data programmatically using the R
programming language along with several add-on
packages \parencite{Grolemund:2011aa,Wickham:2011aa,Couture-Beil:2014aa,Lang:2014aa,R-Core-Team:2014aa}.\footnote{In
addition, the code used to run the analysis, as well as the final
dataset are available as a github repository: \url{https://github.com/pschmied/bikecounts-mle}}
Bicycle counts were aggregated by day, and then joined to weather data
by date. In addition to the variables collected from these two
sources, we were also interested in controlling for holidays and
whether or not the nearby University of Washington was in session.
These data were collected and coded manually from the University of
Washington's historic academic calendars.

Figures \ref{fg:hist} and \ref{fg:timeseries} provide visual summaries
of the processed counts data. Some apparent outliers are visible at
the rightmost portion of the histogram. The two highest counts
occurred the Monday and Tuesday preceding the beginning of National
Bike to Work Month. And the third highest count occurred on National
Bike to Work Day.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{../fig/dp1}
  \caption{Histogram showing frequency of observed bicycles}
  \label{fg:hist}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{../fig/dp2}
  \caption{Timeseries plot of bicycle counts}
  \label{fg:timeseries}
\end{figure}

\FloatBarrier
\subsection*{Variable selection}
% 
% Data analysis: specification of model, choice of model (nb v pois),
% description of counterfactual simulation using the estimated model,
% software used.
%

Out of the set independent variables retrieved from Forecast.io and
the University of Washington, we selected a subset that we felt best
reflected our specific research questions. The final set of variables
included in our model include Table \ref{tb:variables} summarizes the
variables used in our model.

\begin{centering}
\begin{table}[ht!]
\begin{scriptsize}
\caption{Variables included in model specification}
\begin{tabularx}{0.47\textwidth}{>{\raggedright\arraybackslash}p{1.6cm}>{\raggedright\arraybackslash}X}
\toprule
Variable & Description \\
\midrule
Count* & Number of bicycles per day \\
Daylight  & Length of daylight in hours \\
UW & The University of Washington was in session (TRUE/FALSE) \\
Holiday & The day was a holiday as recognized by UW (TRUE/FALSE) \\
Max temp & Maximum temperature in Fahrenheit for the day \\
Max precip & Daily maximum inches of precipitation in any hour \\
Sat--Fri & Day of the week dummy variable (TRUE/FALSE, relative to Sunday) \\
Day \# & Sequentially numbered day of study \\
\bottomrule
\multicolumn{2}{r}{* Dependent variable}
\end{tabularx}

\label{tb:variables}
\end{scriptsize}
\end{table}
\end{centering}

Daylight hours (defined as $sunset - sunrise$) and the University of
Washington in-session status were selected to represent seasonality.
Daylight hours was chosen instead of a calendar-based categorization
of season in part because Seattle's Pacific Maritime Climate differs
substantially from traditional notions of Spring, Summer, Autumn, and
Winter. Daylight hours also is measured as a continuous value at a
finer temporal resolution of one day. Finally, daylight hours adjusts
according to latitude, which may make this model estimation procedure
and specification more transferable to other sites in the future,
perhaps by interacting latitude with daylight hours.

We deemed the University of Washington variable important in part
because of the Fremont Bridge's proximity and connection via the Burke
Gilman Trail to the University of Washington. We also felt that this
variable was a suitable proxy for the ``school season,'' which more
broadly captures whether or not other local schools are in session.
The academic calendars of the various local schools do not align
perfectly, however they still overlap substantially with the
University of Washington, which is itself the largest educational
institution in the region.

Inclusion of the holiday variable was an attempt to account for
some low outlier counts. Upon inspection of the dataset, Christmas and
Thanksgiving in particular had very low counts of bicycles relative to
the days preceding and following. Relatedly, but not accounted for by
any variable in our model, are some of the high outlier counts. Upon
inspection, some of the highest counts were observed on National Bike
to Work Day and on the day of the Fremont Solstice Parade, which
typically draws large numbers of bicyclists as participants and
spectators. The omission of such a variable is justified based on the
few occurrences of high outlier counts, and our desire for this model
to only include variables that could be collected or straightforwardly
adapted to other locations.

Daily maximum temperature, measured in Fahrenheit, was chosen to
represent temperature (rather than, for example, substituting or
adding daily minimum temperature) in part to retain simplicity in the
model, in part because there is relatively little daily temperature
variation in Seattle due to the moderating effect of large water
bodies, and in part because maximum temperature better reflects the
conditions during daylight hours when most bicycle trips would occur.
This simplification may not be warranted for other locations that
experience greater temperature variation than Seattle.

Maximum precipitation, which measures the maximum inches of
precipitation that occurred in any hour throughout the day, was chosen
rather than average precipitation based on the notion that bicyclists
might make travel decisions based on a likely worst case scenario.
This assumption is slightly more problematic than our assumptions
about temperature, in that we do expect bicyclists to be at least
somewhat sensitive to average conditions or conditions observed at
their time of departure. As in the case of temperature, this
simplification would be less justifiable in locations that experience
greater daily variation in precipitation or in locations that have a
predictable pattern of precipitation during certain hours.

Day of the week was added due to its presence in the literature, as
well as an apparent weekly pattern revealed visually by zooming into
the timeseries plot. These data were coded as a set of Boolean dummy
variables, excluding Sunday as the reference category.

The final variable, the day number, was included so that we could test
for a linear trend in bicycling volumes. We created this variable by
sequentially numbering (1--720) the observed counts by day during the
study period.

\FloatBarrier
\subsection*{Model estimation and goodness of fit}
% Model choice
A natural model choice for count data is the Poisson model, however we
believed that our count data were overdispersed. Overdispersion, or
contagion between events, violates the mean-variance equivalence
assumption of the Poisson model. While overdispersion would not impact
our parameter estimates, it would result in overly optimistic margins
of error. In order to account for and estimate the amount of
overdispersion present in our data, we chose a negative binomial model
type. Because there were no days observed with zero bicycle counts, we
did not need to resort to zero-inflated models as is often necessary
when dealing with count data.

% Estimation method
We fit the model in R using the \texttt{glm.nb} function from the MASS
package \parencite{Venables:2002aa}. For comparison, we also estimated
a Poisson model with an analogous specification in R using the
\texttt{glm} function. A much lower AIC and BIC in the case of the
negative binomial model confirmed that it was a better fit than the
Poisson. We then tested for overdispersion using the \texttt{odTest}
function from the Pscl package \parencite{Jackman:2014aa}. The highly
significant chi-squared test statistic provided strong evidence that
overdispersion was present in the data.

Additionally, we compared the residuals plots for both the Poisson and
negative binomial models. As seen in figures \ref{fg:resnb} and
\ref{fg:respois}, the residual variance is much lower in the negative
binomial model, further confirming the choice of a negative
binomial for these counts.

As a final model comparison, we visually assessed the overall fit of
our negative binomial model by plotting actual versus predicted values
as shown in figure \ref{fg:avp}. This fit appears to be generally
good, though very high count days are predicted somewhat less
accurately.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{../fig/residuals_pois}
  \caption{Residual variance from the Poisson model}
  \label{fg:respois}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{../fig/residuals}
  \caption{Residual variance from the negative binomial model}
  \label{fg:resnb}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{../fig/avpp2-sm}
  \caption{Actual versus predicted values as fit by negative binomial model}
  \label{fg:avp}
\end{figure}

% Model interpretation
In order to provide results that are more readily interpretable by
non-statisticians, we used counterfactual simulation to isolate
individual terms from the model that correspond to our research
questions. In so doing, we simulated various quantities of interest
including point estimates and confidence intervals, and then plotted
them for visual inspection. Counterfactual simulations were performed
with a modified version of the Simcf R package, and visualized with
ggplot2 \parencite{Adolph:2014aa,Schmiedeskamp:aa,Wickham:2009aa}.
Results of these simulations are presented in the following section.

\section*{Results}
% 
% One subsection for each research question
% 
This section presents the results from the statistical model and
accompanying counterfactual simulations as described in the preceding
section. Each of the main research questions of seasonality, weather,
and general trend are addressed here. In addition, additional added
control variables such as day of the week, holidays, and linear trend
are presented.

Without exception, each of the coefficients in our model was
statistically significant at the $p < 0.05$ level. Further, with the
exception of the Saturday coefficient, all coefficients were
significant at the $p < 0.001$ level. The remainder of this section
focuses on presenting the substantive effect of each variable.

\FloatBarrier
\subsection*{Seasonality}
As discussed previously, we considered two variables to address the
question of seasonality: the first being the number of daylight hours,
and the second being whether or not the University of Washington
(proxying more generally for other educational institutions) was in
session.

Figure \ref{fg:seasonality} shows that we see a substantial increase
in bicycle volumes when the University of Washington was in session.
With all other factors held constant we see that, on days when the
university is in session, we would expect an average of approximately
367 additional bicycles observed. Similarly, we see a roughly linear
increase in bicycles with increased day length.

\begin{figure}[htbp!]
  \centering
  \includegraphics[width=.5\textwidth]{../fig/m1c4}
  \caption{Effect of daylight hours and University of Washington
    in-session status on bicycle counts, with shaded 95\% confidence
    regions.}
  \label{fg:seasonality}
\end{figure}

\FloatBarrier
\subsection*{Weather}
As in the case of seasonality, our model represents weather using two
variables, in this case precipitation (measured as the maximum amount
of precipitation falling in any hour of that day), and maximum
temperature (measured daily, in Fahrenheit).

The effect of precipitation is shown in figure \ref{fg:precipitation}.
From this, we can see a clear inverse relationship between
precipitation and bicycle counts. The rate of decrease in bicycles
appears to begin somewhat steeply, and then begins to slow slightly at
higher amounts of precipitation. This suggests that people are
generally more sensitive to the presence of precipitation than the
intensity.

Temperature, in contrast to precipitation, has a clearly positive
association with increased numbers of bicyclists. We were somewhat
surprised to not see a leveling off in counts at very high
temperatures---likely due to Seattle's moderate summer climate, and the
lack of extremely warm days in our dataset.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.5\textwidth]{../fig/m1c2}
  \caption{Effect of precipitation on counts, with shaded 95\%
    confidence region.}
  \label{fg:precipitation}
\end{figure}

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.5\textwidth]{../fig/m1c1}
  \caption{Effect of temperature on counts, with shaded 95\%
    confidence region.}
  \label{fg:temperature}
\end{figure}

\FloatBarrier
\subsection*{General trend in bicycle counts}
Our results (shown in figure \ref{fg:trend}) confirm the presence of a
general trend toward increased numbers of bicycles at this location.
With all else held constant, we would expect to see roughly 300 more
bicycles on days at the end of our study period than at the beginning.
This is consistent with figures reported elsewhere that suggest that
bicycling is increasing in a number of cities including
Seattle \parencite{League-of-American-Bicyclists:aa}.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.5\textwidth]{../fig/m1c5}
  \caption{General trend in bicycling counts, all other factors held
    constant, with 95\% confidence region.}
  \label{fg:trend}
\end{figure}

\FloatBarrier
\subsection*{Day of week variation}
As discussed previously, day of the week was included due to the
weekly variation in bicycle counts apparent in timeseries plots.
Figure \ref{fg:dayofweek} shows several interesting aspects of these
results. First, we see much higher numbers of bicyclists on weekdays
than on weekends. This strongly suggests that the majority of the
bicycle traffic at this location is for commuter purposes.

Comparing between weekday results, we see that most days are roughly
the same, with the some drop-off toward Thursday and Friday. This
decrease toward the end of the work week might be attributable to
individuals either working non-traditional schedules or perhaps people
adjusting their travel mode choice in order to accommodate social
engagements.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.5\textwidth]{../fig/m1c3}
  \caption{Variation in counts throughout the week, with 95\% confidence bars.}
  \label{fg:dayofweek}
\end{figure}

\FloatBarrier
\section*{Discussion}
This research set out to better understand the relationship between
bicycle counts and weather and season. However, weather and season are
themselves comprised by a number of constituent elements. We believe
this research has chosen defensible representations of weather and
season. However, we readily concede that there exist other reasonable
representations of season and weather.

In particular, others might be more interested in building a model
that places more emphasis on predictive power. Those individuals might
choose to add a number of additional variables or consider more
complex interactions between terms. By contrast, in this model
formulation, we present a relatively simple model with comparatively
fewer terms. This was due to our interest in a conceptual
understanding of the influence of weather and season.

Even within our model, as noted in the methodology section, we
considered alternative specifications, especially those containing
different measures of temperature and precipitation. As an example, in
choosing to represent temperature as the daily maximum, we make some
assumptions about people and their mode choice decision making
processes. People cannot know the daily maximum temperature in advance
of their morning commute, however the measured daily maximum
temperature is likely to be similar to temperatures reported in
temperature forecasts. A superior measure might be a consensus of
forecasts available in the morning of each day. Similarly, if
Seattle's daily high and daily low temperatures did not have such low
variation, a daily maximum temperature might be less important to a
winter-time bicyclist than daily minimum.

One variable missing from our analysis that we would have liked to
include is cloud cover.\footnote{We gratefully acknowledge Christopher
  Adolph for this suggestion.} This variable would have allowed us to
account for the possibility of brilliant, sunny, but cold winter days,
as well as warm, dry, but gloomy summer days. Unfortunately, the
dataset downloaded from Forecast.io included a great deal of missing
values for this variable, and we were unable to identify another
suitable source for this data in time for this article.

Another limitation of this research comes from our sample of bicycle
counts. While we do have continuous counts spanning a full two years,
these counts were taken at just one point location. We believe the
Fremont bridge is somewhat representative of a high-volume bicycle
facility in Seattle, however we might, for example, see a decreased
effect size for University of Washington session status if we looked
at counts further from the university. The City of Seattle has begun
collecting and releasing data from other counters, which could be
included in a future version of this analysis.

Finally, as discussed in the methodology section, we believe the
choice of the negative binomial model to be a better choice than the
Poisson, which had been used in some previous studies. A limitation in
this model, however, is in not accounting for timeseries
autocorrelation. We think it highly likely that the decision to
bicycle on one day might influence the decision to bicycle on the
next. A future direction would be to consider models, such as ARIMA,
that more explicitly account for time.

\section*{Conclusion}
This research set out to help clarify the relationship between weather
and seasonal factors and bicycle counts in Seattle. In order to
achieve this, we developed a negative binomial model to predict
bicycle counts based on temperature, precipitation, day length,
university in-session status, day of the week, and a general linear
time trend.

For each term included in our model, we found statistically
significant effects. More importantly, through the use of
counterfactual simulation, we estimated what we deem to be substantial
effect sizes associated with each predictor variable.

This article contributes to the existing literature by demonstrating
the use of a negative binomial model for bicycle counts. In addition,
the results presented here were generated from data collected over a
long study period of two years.

While control of the weather and seasons are admittedly beyond the
scope of policy makers, this research does suggest that planners and
policy makers may want to develop strategies that help mitigate the
impacts of the natural environment during the winter months on the
increasing number of bicyclists. In other words, the delta between
warm dry days and cold wet days should be treated as the opportunity
frontier. Future research could focus on determining what, if any,
programmatic or built interventions could ameliorate unfavorable cold-
and wet-weather bicycling conditions.

\printbibliography
\end{document}